---
title: "Adult survival scenarios"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adult survival scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TACAR)
```

Establish distribution of plausible adult survival values in the 
following scenarios: 

 - Unharvested, with and without recovery actions.
 - Lightly harvested, with and without recovery actions.
 - Intesively harvested, with and without recovery actions.
 
 To be able to compare with and without recovery actions we need to 
 include density dependence in the matrix projection models.
 
 We use estimates of longevity to establish mean and range of annual 
 adult female survival in each scenario.

 - Unharvested,    
 mean 57.1 years, range 20 to 70 years, beta distribution.
 
 - Lightly harvested (less than 2.5% per year),    
 mean 30 years, range 20 to 70 years, beta distribution.
 
 - Intensively harvested (greater than 5% per year),    
 mean 22 years, range 20 to 70 years, half-normal distribution.
 
```{r survival-distributions}
adult_survival <- TACAR::make_beta(survival_mean = 0.976987, 
                                      survival_min = 0.9215, 
                                      survival_max = 0.98153, 
                                   alpha = 40, beta = 40
                                   )
hist(adult_survival$within_range)
mean(adult_survival$within_range, na.rm = TRUE)

adult_survival_hunt_low <- TACAR::make_beta(survival_mean = 0.95244, 
                                      survival_min = 0.9215, 
                                      survival_max = 0.98153, 
                                   alpha = 400, beta = 400
                                   )
hist(adult_survival_hunt_low$within_range)
mean(adult_survival_hunt_low$within_range, na.rm = TRUE)

adult_survival_hunt_high <- TACAR::make_half_normal(survival_mean = 0.96,
                                                    survival_min = 0.92, 
                                                    survival_max = 0.97339)
# Plot the distribution
hist(adult_survival_hunt_high$survival_values, probability = TRUE)
mean(adult_survival_hunt_high$survival_values, na.rm = TRUE)


```
 

