---
title: "Adult survival scenarios"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adult survival scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(TACAR)
library(ggplot2)
library(purrr)
library(tibble)
library(popdemo)
library(popbio)
library(Rcompadre)
```

## Population scenarios

Establish distribution of plausible adult survival values in the 
following scenarios: 

 - Unharvested, with and without recovery actions.
 - Lightly harvested, with and without recovery actions.
 - Intesively harvested, with and without recovery actions.
 
 The "unharvested" scenario represents the baseline reference against 
 which the others will be measured. To be able to compare with and 
 without recovery actions we need to include density dependence 
 in the matrix projection models.
 
 We use estimates of longevity to establish mean and range of annual 
 adult female survival in populations for each scenario. 
 This is then used to estimate the distribution  "today", 
 before considering the effect of future actions (either positive e.g. 
 recovery or negative e.g. continued harvest).

 - Unharvested ("none"),    
 mean 57.1 years, range 20 to 70 years, beta distribution.
 
 - Lightly harvested ("low" - less than 5% per year),    
mean 22 years, range 12.2 (9 plus 3 nesting seasons) to 50 years, beta distribution.
 
 - Intensively harvested ("high" - greater than 10% per year),    
 mean 12.2 years, range 9 (reproduces at least once) to 30 years, half-normal distribution.

```{r adult-survival, echo=FALSE}
# Captive logevity ranges between 20 to 70 years.
# Captive survival ranges between 0.9215 - 0.98153

# Anecdotal evidence suggests they can live up to 70 years 
# (Ricklefs 2010; Myhrvold et al. 2015; AnAGe 2023 - 
#online webserver https://genomics.senescence.info/species/index.html). 
# In captivity, individuals have been reported to live from 23.4 to 70.
# Using Iverson’s (in press) regression equation of longevity 
# [L = 2.741*(age at maturity) + 37.926] and the range of values for 
# female age at first reproduction (5 - 9 years), 
# the mean L was estimated at 57.11 years and to range between 51.63 
# and 62.60 years, which is relatively close to the anecdotal evidence.

# Get matrix
# projection parameters
dt <- TACAR::pop01_param_poun()
# Population size constant.
nf <- 10
dt$adultF_n <- nf
# Matrix from data.frame
stage_names <- c("a1", "a2", "a3", "a4",
                   "b1", "b2", "b3", "b4",
                   "c1", "c2", "c3", "c4",
                   "d1", "d2", "d3", "d4")
vpop <- unlist(dt[3 , stage_names])
pop_mat <- matrix(vpop, byrow = TRUE, ncol=4)
dimnames(pop_mat) <- list(c("_1", "_2", "_3", "_4"),
                            c( "a", "b", "c", "d"))

# Data with life expectancy, female survival and lambda
# Mpm with different female survival
suvr_f <- seq(0.5, 0.99, by = 0.0001)
ntimes <- length(suvr_f)
pop_mat_s <- rep(list(pop_mat), ntimes)
pop_mat_s <- mapply(function(A, r){A[4, 4] <- r; A}, pop_mat_s, suvr_f, 
                      SIMPLIFY = FALSE)
# Function to get life expectancy and lambda
calculate_le_mean_tidy <- function(pop_matrix) {
mat_u <- pop_matrix
mat_u[1, 4] <- 0
tibble::tibble(life_expectancy_mean =  sum(Rage::life_expect_mean(matU = mat_u, 
                                                start = NULL)), 
               lambda = round(popbio::lambda(pop_matrix), 6))
}

# Use purrr::map to apply the function to each population_matrix
survival_longevity <- purrr::map_dfr(pop_mat_s, calculate_le_mean_tidy)
survival_longevity$female_survival <- suvr_f

```

 
```{r survival-distributions}
adult_survival <- TACAR::make_beta(survival_mean = 0.976987, 
                                      survival_min = 0.9215, 
                                      survival_max = 0.98153, 
                                   alpha = 40, beta = 40
                                   )
hist(adult_survival$within_range, probability = TRUE)
mean(adult_survival$within_range, na.rm = TRUE)

adult_survival_hunt_low <- TACAR::make_beta(survival_mean = 0.9326, 
                                      survival_min = 0.8407, 
                                      survival_max = 0.97339, 
                                   alpha = 40, beta = 30
                                   )
hist(adult_survival_hunt_low$within_range, probability = TRUE)
mean(adult_survival_hunt_low$within_range, na.rm = TRUE)

adult_survival_hunt_high <- TACAR::make_half_normal(survival_mean = 0.8407,
                                                    survival_min = 0.7363, 
                                                    survival_max = 0.9525)
# Plot the distribution
hist(adult_survival_hunt_high$survival_values, probability = TRUE)
mean(adult_survival_hunt_high$survival_values, na.rm = TRUE)

```

Compare the survival in populations under different hunting pressure.

```{r plot-density, message=FALSE, warning=FALSE}
df_survival <- rbind(
data.frame(distribution = "beta", hunt_level = "a_none", 
           female_survival = adult_survival$within_range),
data.frame(distribution = "beta", hunt_level = "b_low", 
           female_survival = adult_survival_hunt_low$within_range),
data.frame(distribution = "half-normal", hunt_level = "c_high", 
           female_survival = adult_survival_hunt_high$survival_values) 
)
df_survival$hunt_levelf <- factor(df_survival$hunt_level)
levels(df_survival$hunt_levelf) <- c("none", "low", "high")

#
growth_survival <- survival_longevity[which.min(abs(1-survival_longevity$lambda)), ] |> 
  dplyr::pull(female_survival)

# Plot
mylabels <- c("0.75\n(9.5 Y)", "0.85\n(12.7 Y)", "0.95\n(28.8 Y)")
ggplot(data = df_survival, aes(x = female_survival, 
                               group = hunt_levelf, fill = hunt_levelf)) + 
  geom_density(alpha = 0.5, adjust = 2) + 
   geom_vline(aes(xintercept = growth_survival), 
    colour = "blue") + 
  #annotate("label", x = growth_survival, y = 25, label = "Some text") +
  scale_fill_viridis_d("harvest") + 
  scale_x_continuous(breaks = c(0.75, 0.85, 0.95), 
                     labels = mylabels) +
  xlab("Female survival") +
  ylab("Density")

```

The figure above shows the distribution of adult female survival in 
populations under threee different scenarios. The x axis shows 
survival values, together with overall life expectancy in parentheses. 
The vertical blue line shows where lambda equals one, and indicates 
where populations are growing to the right.

 
We can now use these distributions to explore how the populations are 
likely to change in the future. Here we follow the element selection 
example: 
https://cran.r-project.org/web/packages/popdemo/vignettes/popdemo.html#72_Nonrandom_matrix_selection

First run an example using the Intensively harvested population, 
with harvest at 13% a year.
```{r project-survival}
# Get matrix
# projection parameters
dt <- TACAR::pop01_param_poun()
# Population size constant.
nf <- 10
dt$adultF_n <- nf

# No data on density dependence for any South American freshwater turtle.
# Here include a generous 20% ceiling on maximum population increase 
# in the case of harvested populations.
ceiling_threshold_high <- ceiling(nf + (nf * 0.2))
ceiling_threshold_low <- ceiling(nf + (nf * 0.1))
ceiling_threshold_none <- ceiling(nf + (nf * 0.05))

# Matrix from data.frame
stage_names <- c("a1", "a2", "a3", "a4",
                   "b1", "b2", "b3", "b4",
                   "c1", "c2", "c3", "c4",
                   "d1", "d2", "d3", "d4")
vpop <- unlist(dt[3 , stage_names])
pop_mat <- matrix(vpop, byrow = TRUE, ncol=4)
dimnames(pop_mat) <- list(c("_1", "_2", "_3", "_4"),
                            c( "a", "b", "c", "d"))
  # stable stage structure
  pop_ss <- popdemo::eigs(pop_mat, "ss")
  # stable stage population numbers corresponding to number of females
  pop_n <-  nf * (pop_ss * (1/pop_ss[4]))
#create list of 100 matrices
ntimes <- 100
pop_mat_100 <- rep(list(pop_mat), ntimes)
#replace [d, 4] element of every matrix with new random number
a44 <- sample(na.omit(adult_survival_hunt_high$survival_values), 
              size = ntimes)
pop_mat_100 <- mapply(function(A, r){A[4, 4] <- r; A}, pop_mat_100, a44, 
                      SIMPLIFY = FALSE)
# Deterministic
# Average difference (0.976987 - 0.8407) / 0.976987 = 0.139
pop_mat_high <- pop_mat
pop_mat_high[4, 4] <- (pop_mat[4, 4] - (pop_mat[4, 4]*0.139))
# Project and plot
poun_high <- popdemo::project(pop_mat_100, vector = pop_n, Aseq = 1:ntimes)
plot(poun_high)
lines(0:ntimes, project(pop_mat_high, pop_n, time = ntimes), 
      lty = 3, col = "red")
```

With continued harvest and no conservation actions the population 
declines rapidly.
We now repeat the process and compare future changes for the three 
different scenarios.
```{r compare-scenarios}
# No hunting
pop_mat_100_none <- rep(list(pop_mat), ntimes)
#replace [d, 4] element of every matrix with new random number
a44_none <- sample(na.omit(adult_survival$within_range), size = ntimes)
# Update values
pop_mat_100_none <- mapply(function(A, r){A[4, 4] <- r; A}, 
                           pop_mat_100_none, a44_none, 
                      SIMPLIFY = FALSE)
# Project
poun_none <- popdemo::project(pop_mat_100_none, 
                              vector = pop_n, Aseq = 1:ntimes)

# Low hunting
pop_mat_100_low <- rep(list(pop_mat), ntimes)
#replace [d, 4] element of every matrix with new random number
a44_low <- sample(na.omit(adult_survival_hunt_low$within_range), 
                  size = ntimes)
# No do not add additional "Perturbation" of 2.5% harvest
# a44_low <- (a44_low - (a44_low * 0.025))
# Update values
pop_mat_100_low <- mapply(function(A, r){A[4, 4] <- r; A}, 
                           pop_mat_100_low, a44_low, 
                      SIMPLIFY = FALSE)
# Project
poun_low <- popdemo::project(pop_mat_100_low, 
                              vector = pop_n, Aseq = 1:ntimes)

# Data for plotting-Adult females at each time point for each scenario.
len <- length(poun_none)
Time.intervals <- 0:(len - 1)
adult_fe_none <- as.integer(trunc(poun_none * (popbio::stable.stage(pop_mat)[4])))
adult_fe_low <- as.integer(trunc(poun_low * (popbio::stable.stage(pop_mat)[4])))
adult_fe_high <- as.integer(trunc(poun_high * (popbio::stable.stage(pop_mat)[4])))

df_adult_female <- rbind(
data.frame(year = Time.intervals, hunt_level = "a_none", 
           females = adult_fe_none),
data.frame(year = Time.intervals, hunt_level = "b_low", 
           females = adult_fe_low),
data.frame(year = Time.intervals, hunt_level = "c_high", 
           females = adult_fe_high) 
)
df_adult_female$hunt_levelf <- factor(df_adult_female$hunt_level)
levels(df_adult_female$hunt_levelf) <- c("none", "low", "high")

# Plot
ggplot(data = df_adult_female, aes(x = year, y = females, 
                                   group = hunt_levelf, 
                                   colour = hunt_levelf)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  scale_y_continuous(transform = "log1p") +
  scale_colour_viridis_d("harvest")
```

These models are useful but could be more realisitic.
The figure above shows that with no or little hunting the population 
numbers increase at such an unrealistic rate that we need to use a 
log scale to compare with the intensively harvested population. 

For example, to enable us to explore potential future impacts of 
conservation actions it is important to consider how density 
dependence will affect the survival and transient dynamics of 
recovering populations. This will enable more robust estimation 
of expected outcomes and answer questions such as how long will 
recovery take and how much will recovery cost?


## Density dependence

Working on it .........

Examples using Chrissy´s new function for a threshold-based approach.

### Fertility/fecundity

```{r threshold-dd}
# Under no hunting, we would follow pop_mat_100_none.
pop_mat_nohunt<- pop_mat_100_none[[1]]
pop_mat_nohunt[4,4]<- mean(a44_none)

# For a high-density population, let's say that they would have a lambda value
# of 1 on average. We can knock down the population growth rate to 1 by dividing the
# fertility entries by R0:
# No - high density populations should decline.
R0<- popbio::net.reproductive.rate(pop_mat_nohunt)
nohunt_highD<- pop_mat_nohunt
nohunt_highD[1,4]<- nohunt_highD[1,4]/R0
# Check the lambda value:
Re(eigen(nohunt_highD, only.values = TRUE)$values[1])
popbio::lambda(nohunt_highD)
# Try out the DD projection function when we use a single matrix:
nohunt_threshold<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=1000, A_DD=nohunt_highD)
nohunt_threshold_500<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=500, A_DD=nohunt_highD)
#population can double to 310 before DD kicks in - why do females increase fourfold?
nohunt_threshold_x2<- project_threshold(pop_mat_100_none, 
                                         vector=pop_n, time=100, 
                                         threshold=ceiling(sum(pop_n)*2), 
                                         A_DD=nohunt_highD)
# Plot the result:
data_adults<- nohunt_threshold$vec[,4]
data_adults_500<- nohunt_threshold_500$vec[,4] 
data_adults_x2<- nohunt_threshold_x2$vec[,4] 
plot(data_adults, type='l', lwd=2, xlab='Time', ylab='N adult females')
lines(0:100, data_adults_500, lwd=2, col = "red")
lines(0:100, data_adults_x2, lwd=2, col = "blue")

# If we want to keep the stochasticity, but we want to drop the reproductive
# output when the population is above the threshold size, we can divide the
# reproductive output in the whole series of population matrices by R0:
pop_mat_100_none_highdens<- lapply(pop_mat_100_none, function(x){x[1,4]<- x[1,4]/R0; return(x)})

# Try out the DD projection function when we still have stochasticity:
nohunt_threshold_stoch<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=1000, A_DD=pop_mat_100_none_highdens)
nohunt_threshold_stoch_500<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=500, A_DD=pop_mat_100_none_highdens)
#population can double to 310 before DD kicks in - why do females increase fourfold?
nohunt_threshold_stoch_x2<- project_threshold(pop_mat_100_none, 
                                              vector=pop_n, time=100, 
                                              threshold=ceiling(sum(pop_n)*2), 
                                              A_DD=pop_mat_100_none_highdens)
# Plot the result:
data_adults_stoch<- nohunt_threshold_stoch$vec[,4]
data_adults_stoch_500<- nohunt_threshold_stoch_500$vec[,4]
data_adults_stoch_x2<- nohunt_threshold_stoch_x2$vec[,4]
plot(data_adults_stoch, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')
lines(0:100, data_adults_stoch_500, lwd=2, col = "red")
lines(0:100, data_adults_stoch_x2, lwd=2, col = "blue")

# For populations that do not experience hunting - a threshold of 20%
# seems reasonable.
# Number of adult females should stay around 10 - 12
# Mean matrix
pop_mat_mean_none <- Rcompadre::mat_mean(pop_mat_100_none)
# stable stage structure
pop_ss_none <- popdemo::eigs(pop_mat_mean_none, "ss")
# stable stage population numbers corresponding to number of females
nf_thresh_none <- nf*1.1 # 10% increase plausible
pop_none <-  nf_thresh_none * (pop_ss_none * (1/pop_ss_none[4]))
pop_thresh_none <- ceiling(sum(pop_none)) #164

# Using this approach the population triples.
# Long lived species where adult survival is most important.
# DD changes to fertility take a very long time to show.
nohunt_t10_stoch<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n, time=100, 
                                     threshold=pop_thresh_none, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_stoch10 <- nohunt_t10_stoch$vec[,4]
plot(data_adults_stoch10, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')

# Now use simple threshold.
nohunt_t10_stoch2<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n, time=100, 
                                     threshold=11, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_t10_stoch2 <- nohunt_t10_stoch2$vec[,4]
plot(data_adults_t10_stoch2, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')

summary(data_adults_t10_stoch2)

# Check pattern not a rtesult of inital pop size.
pop_n100 <- pop_n*10
nohunt_t100_stoch2<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n100, time=100, 
                                     threshold=110, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_t100_stoch2 <- nohunt_t100_stoch2$vec[,4]
plot(data_adults_t100_stoch2, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')

summary(data_adults_t100_stoch2)

pop_n1000 <- pop_n*100
nohunt_t1000_stoch2<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n1000, time=100, 
                                     threshold=1100, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_t1000_stoch2 <- nohunt_t1000_stoch2$vec[,4]
plot(data_adults_t1000_stoch2, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')

summary(data_adults_t1000_stoch2)

# What happens if we reduce threshold to very low level?
nohunt_t9_stoch2<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n, time=100, 
                                     threshold=9, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_t9_stoch2 <- nohunt_t9_stoch2$vec[,4]
plot(data_adults_t9_stoch2, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')
summary(data_adults_t9_stoch2)

nohunt_t5_stoch2<- project_threshold(pop_mat_100_none, 
                                     vector=pop_n, time=100, 
                                     threshold=0, 
                                     A_DD=pop_mat_100_none_highdens)
data_adults_t5_stoch2 <- nohunt_t5_stoch2$vec[,4]
plot(data_adults_t5_stoch2, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')
summary(data_adults_t5_stoch2)

# See how number of females changes with different thresholds.

# wrap function 
test_thresh <- function(x){
mythresh <- x  
proj_res<- project_threshold(pop_mat_100_none, 
                         vector=pop_n, time=100, 
                         threshold=mythresh, 
                         A_DD=pop_mat_100_none_highdens)
data_females <- proj_res$vec[,4]
meanf <- mean(data_females)
myres <- data.frame(threshold = mythresh, mean_females = round(meanf, 2))
myres
}
tseq <- 0:1000
dft <- purrr::map_dfr(tseq, test_thresh)
ggplot(dft, aes(x = threshold, y = mean_females)) +
  geom_point()+ 
  stat_smooth()

```


### Fertility/fecundity and female survival

Now include mortality so population growth is less than 1 - 
high density populations decline.

Increased population density generally acts to increase mortality 
rates and/or reduce survival probabilities across a diverse spectrum 
of wild vertebrate species. While the specific manifestations vary, 
the overarching pattern is one of negative feedback, where denser 
populations experience greater per capita demographic costs.

Here a 10% loss in annual female survival (increased competition/disease/predation).

```{r threshold-dd-survival}
# Use nohunt
nohunt_highDs<- nohunt_highD
nohunt_highDs[4,4]<- nohunt_highD[4,4] - (nohunt_highD[4,4]*0.1)
# Check the lambda value:
Re(eigen(nohunt_highDs, only.values = TRUE)$values[1])
popbio::lambda(nohunt_highDs)

# Try out the DD projection function when we use a single matrix:
nohunt_thresholdDs<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=1000, A_DD=nohunt_highDs)
nohunt_thresholdDs_500<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=500, A_DD=nohunt_highDs)
#population can double to 310 before DD kicks in - why do females increase fourfold?
nohunt_thresholdDs_x2<- project_threshold(pop_mat_100_none, 
                                         vector=pop_n, time=100, 
                                         threshold=ceiling(sum(pop_n)*2), 
                                         A_DD=nohunt_highDs)
# Plot the result:
data_adultsDs<- nohunt_thresholdDs$vec[,4]
data_adultsDs_500<- nohunt_thresholdDs_500$vec[,4] 
data_adultsDs_x2<- nohunt_thresholdDs_x2$vec[,4] 
plot(data_adultsDs, type='l', lwd=2, xlab='Time', ylab='N adult females')
lines(0:100, data_adultsDs_500, lwd=2, col = "red")
lines(0:100, data_adultsDs_x2, lwd=2, col = "blue")

# Now same with stochastic
pop_mat_100_none_highdensDs<- lapply(pop_mat_100_none_highdens, 
                                     function(x){x[4,4]<- x[4,4]-(x[4,4]*0.1); return(x)})

# Try out the DD projection function when we still have stochasticity:
nohunt_threshold_stochDs<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=1000, A_DD=pop_mat_100_none_highdensDs)
nohunt_threshold_stochDs_500<- project_threshold(pop_mat_100_none, vector=pop_n, time=100, threshold=500, A_DD=pop_mat_100_none_highdensDs)
nohunt_threshold_stochDs_x2<- project_threshold(pop_mat_100_none, 
                                              vector=pop_n, time=100, 
                                              threshold=ceiling(sum(pop_n)*2), 
                                              A_DD=pop_mat_100_none_highdensDs)
# Plot the result:
data_adults_stochDs<- nohunt_threshold_stochDs$vec[,4]
data_adults_stochDs_500<- nohunt_threshold_stochDs_500$vec[,4]
data_adults_stochDs_x2<- nohunt_threshold_stochDs_x2$vec[,4]
plot(data_adults_stochDs, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')
lines(0:100, data_adults_stochDs_500, lwd=2, col = "red")
lines(0:100, data_adults_stochDs_x2, lwd=2, col = "blue")

# See how number of females changes with different thresholds.
# wrap function 
test_threshDs <- function(x){
mythresh <- x  
proj_res<- project_threshold(pop_mat_100_none, 
                         vector=pop_n, time=100, 
                         threshold=mythresh, 
                         A_DD=pop_mat_100_none_highdensDs)
data_females <- proj_res$vec[,4]
meanf <- mean(data_females)
myres <- data.frame(threshold = mythresh, mean_females = round(meanf, 2))
myres
}
tseq <- 0:1000
dftDs <- purrr::map_dfr(tseq, test_threshDs)
ggplot(dftDs, aes(x = threshold, y = mean_females)) +
  geom_point()+ 
  stat_smooth()
```

Compare number of females using DD with and without female survival.

```{r compare-dd}

dfc <- rbind(data.frame(dft, dd = "fertility/fecundity"), 
      data.frame(dftDs, dd = "fertility/fecundity\nfemale survival")) 
ggplot(dfc, aes(x = threshold, y = mean_females)) +
  geom_point()+ 
  facet_wrap(~dd)
```

The relationships shown above can be used to select thresholds that
correspond to a stable population with the desired number of 
adult females.

For example, for the case where adult females may increase in the wild 
by 10%, the following code projects a stochastic population.

```{r project-stable}
# threshold for desired number of adults
females_toget <- nf*1.1
thresh_touse <- dftDs[which.min(abs(females_toget-dftDs$mean_females)),] |> 
  dplyr::pull(threshold)
# project stochastic stable population
stable_pop<- project_threshold(pop_mat_100_none, vector=pop_n, 
                               time=100, threshold=thresh_touse, 
                               A_DD=nohunt_highDs)
# Plot the result:
data_adults_stable<- stable_pop$vec[,4]
plot(data_adults_stable, type='l', lwd=2, xlab='Time', 
     ylab='N adult females')

summary(data_adults_stable)

```


## Senesence?

As turtles are long lived species, the populations are highly 
sensitive to adult survival rates (Heppell 1998). Within stage-based 
matrix models this can be better explored by splitting adults 
into three stages:

 - young adult,    
 with lower survival and fecundity due to smaller body sizes.
 
 - mature adults,    
 with greatest survival and fecundity due to body sizes and general health.

 - elderly adults,    
 with lower survival and fecundity due to senescence.
 
 There appears to be considerable variation in senesence in reptiles, 
 and there are few data available that evaluate senesence in long-lived
 turtles (Reinke et al. 2022). But the available evidence suggests 
 reproduction and survival can decline with age (Warner et. al. 2016).
 For example, despite relatively high reproductive output, older 
 females produced eggs with reduced hatching success 
 (Congdon et. al. 2001, Warner et. al. 2016). 
 A study from zoos and aquariums also showed that turtles and 
 tortoises may  reduce senescence in response to improvements in 
 environmental conditions (da Silva et. al. 2022). 
 This raises the possibility that improved environmental management 
 can reduce senesence in freshwater turtles.
 
 There is no data available for *Podocnemis unifilis* on differences in 
 survival/fecundity with age of adult females. So here we use 
 data from other turtle species with similar life history and 
 demographic characteristics. This includes *Podocenmis expansa* - 
 where survival of different adult stages has been documented 
 (Mogollones et. al. 2010). The Matrix Population Models developed in 
 this study and others are freely available from the COMADRE database 
 (https://compadre-db.org/Data/Comadre). Here we use a 
 subset from COMADRE version 4.23.3.1 created on 24 March 2023.
 
```{r load-comadre, eval=TRUE}
# Data for Podocnemididae
# A COM(P)ADRE database ('CompadreDB') object with 2 SPECIES and 2 MATRICES.
comadre_podoc
```
 
 

## References

da Silva, R., Conde, D. A., Baudisch, A., & Colchero, F. 2022. 
Slow and negligible senescence among testudines challenges evolutionary 
theories of senescence. Science, 376 (6600), 1466-1470.
https://doi.org/10.1126/science.abl7811

Congdon, J. D. et. al. 2001. 
Hypotheses of aging in a long-lived vertebrate, Blanding's turtle 
(Emydoidea blandingii). 
Experimental gerontology, 36(4-6), 813-827.
https://doi.org/10.1016/S0531-5565(00)00242-4

Heppell, S. S. 1998. 
Application of life-history theory and 
population model analysis to turtle conservation. 
Copeia, 367-375. https://doi.org/10.2307/1447430

Mogollones SC, Rodríguez DJ, Hernández O, Barreto GR. 2010.    
A demographic study of the arrau turtle (Podocnemis expansa) 
in the Middle Orinoco River, Venezuela. 
Chelonian Conservation and Biology. Jun 1;9(1):79-89.
https://doi.org/10.2744/CCB-0778.1

Reinke et al. 2022. 
Diverse aging rates in ectothermic tetrapods provide insights for the 
evolution of aging and longevity.
Science 376,1459-1466.
https://doi.org/10.1126/science.abm0151

Warner, D. A., Miller, D. A., Bronikowski, A. M., & Janzen, F. J. 2016. 
Decades of field data reveal that turtles senesce in the wild. 
Proceedings of the National Academy of Sciences, 113(23), 6502-6507.
https://doi.org/10.1073/pnas.1600035113
