---
title: "Adult survival scenarios"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adult survival scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TACAR)
library(ggplot2)
```

Establish distribution of plausible adult survival values in the 
following scenarios: 

 - Unharvested, with and without recovery actions.
 - Lightly harvested, with and without recovery actions.
 - Intesively harvested, with and without recovery actions.
 
 To be able to compare with and without recovery actions we need to 
 include density dependence in the matrix projection models.
 
 We use estimates of longevity to establish mean and range of annual 
 adult female survival in populations for each scenario. 
 This is then used to estimate the distribution  "today", 
 before considering the effect of future actions (either positive e.g. 
 recovery or negative e.g. continued harvest).

 - Unharvested,    
 mean 57.1 years, range 20 to 70 years, beta distribution.
 
 - Lightly harvested (less than 2.5% per year),    
 mean 30 years, range 20 to 70 years, beta distribution.
 
 - Intensively harvested (greater than 5% per year),    
 mean 22 years, range 20 to 70 years, half-normal distribution.
 
```{r survival-distributions}
adult_survival <- TACAR::make_beta(survival_mean = 0.976987, 
                                      survival_min = 0.9215, 
                                      survival_max = 0.98153, 
                                   alpha = 40, beta = 40
                                   )
hist(adult_survival$within_range, probability = TRUE)
mean(adult_survival$within_range, na.rm = TRUE)

adult_survival_hunt_low <- TACAR::make_beta(survival_mean = 0.95244, 
                                      survival_min = 0.9215, 
                                      survival_max = 0.98153, 
                                   alpha = 400, beta = 400
                                   )
hist(adult_survival_hunt_low$within_range, probability = TRUE)
mean(adult_survival_hunt_low$within_range, na.rm = TRUE)

adult_survival_hunt_high <- TACAR::make_half_normal(survival_mean = 0.96,
                                                    survival_min = 0.92, 
                                                    survival_max = 0.97339)
# Plot the distribution
hist(adult_survival_hunt_high$survival_values, probability = TRUE)
mean(adult_survival_hunt_high$survival_values, na.rm = TRUE)

```

Compare the survival in populations under different hunting pressure.

```{r plot-density, message=FALSE, warning=FALSE}
df_survival <- rbind(
data.frame(distribution = "beta", hunt_level = "a_none", 
           female_survival = adult_survival$within_range),
data.frame(distribution = "beta", hunt_level = "b_low", 
           female_survival = adult_survival_hunt_low$within_range),
data.frame(distribution = "beta", hunt_level = "c_high", 
           female_survival = adult_survival_hunt_high$survival_values) 
)
df_survival$hunt_levelf <- factor(df_survival$hunt_level)
levels(df_survival$hunt_levelf) <- c("none", "low", "high")

# Plot
mylabels <- c("0.92\n(19.7 Y)", "0.94\n(24.8 Y)", "0.96\n(34.8 Y)", 
              "0.98\n(64.0 Y)")
ggplot(data = df_survival, aes(x = female_survival, 
                               group = hunt_levelf, fill = hunt_levelf)) + 
  geom_density(alpha = 0.5, adjust = 2) + 
  scale_fill_viridis_d("harvest") + 
  scale_x_continuous(breaks = c(0.92, 0.94, 0.96, 0.98), 
                     labels = mylabels) +
  xlab("Female survival") +
  ylab("Density")

```

The figure shows the distribtution of adult female survival in 
populations under threee different scenarios. The x axis shows 
survival values and associated life expectancy in parentheses.

 
We can now use these distributions to explore how the populations are 
likely to change in the future.
