---
title: "Prepare river points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prepare river points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r packages, message=FALSE, warning=FALSE}
library(TACAR)
library(plyr)
library(dplyr)
library(sf)
#library(terra)
library(tidyselect)
library(units)
```

## Make points

### Make function.
Function to make points for each REACH_ID.
```{r point-function}
# function to make points from Free-flowing River reaches
river_points <- function(x){
# Points by reach
require(sf)
require(dplyr)
require(tidyselect)
require(units)
din <- x
row.names(din) <- NULL
crop.l <- sf::st_as_sf(din)
geom_col_name <- attr(crop.l,"sf_column")
tmpdf <- din |> dplyr::select(!tidyselect::all_of(geom_col_name))
# crop.l <- reach_warn_sf
crop.l <- sf::st_cast(sf::st_sfc(sf::st_geometry(crop.l)),"LINESTRING")
crop.l <- sf::st_union(crop.l)
myl <- as.numeric(units::set_units(sum(sf::st_length(crop.l)), km))
crop.l <- sf::st_cast(sf::st_sfc(sf::st_geometry(crop.l)),"LINESTRING")

# Sample line
if(myl < 1){sf.r.segs <- sf::st_line_sample(crop.l, sample = 0) 
} else {sf.r.segs <- sf::st_line_sample(crop.l, density = units::set_units(1, 1/km))}
# Removes empty gemoetries, caused by sampling reaches with broken lines etc.
sf.r.segs <-  sf.r.segs[(is.na(sf::st_dimension(sf.r.segs)) == FALSE ), ]
rm(crop.l)

# Make points
sf.r.segsp <- sf::st_cast(sf::st_sfc(sf::st_geometry(sf.r.segs)),"POINT")
sf.r.segsp <- sf.r.segsp[which(duplicated(st_coordinates(sf.r.segsp ))==FALSE)]
rm(sf.r.segs)
n_rows <- length(sf.r.segsp)
tmpdf <- tmpdf |> dplyr::slice(rep(1:n(), each = n_rows))
row.names(tmpdf) <- NULL
# Return sf
tmpsf <- sf::st_sf(tmpdf, geom = sf.r.segsp)
rm(tmpdf)
tmpsf$aid <- row.names(tmpsf)
tmpsf$reach_id_point <- paste(tmpsf$REACH_ID, tmpsf$aid, sep="_")

dfout <- tmpsf |> data.frame() 
rm(tmpsf)
return(dfout)
}

```

### Process full data
First make files holding free-flowing rivers and basins for future use.
```{r select-ffr, eval=FALSE}
# Crop to basins from Norris et. al. 2019.
basin_in <- system.file("vector\\shape_basin\\amazon_orinoco.shp", 
                        package = "TACAR")
basins <- read_sf(basin_in)
sf_use_s2(FALSE) # to avoid duplicate vertex error
basins_clean <- st_union(st_buffer(basins, 0), by_feature = TRUE) |> 
  group_by(BASIN_NAME, subbasin) |> 
  summarise(area = min(Area_km2)) |> st_cast("MULTIPOLYGON")
basins_clean$BASIN_FLAG <- as.numeric(as.factor(basins_clean$BASIN_NAME))
basins_clean$SUBBASIN_FLAG <- as.numeric(as.factor(basins_clean$subbasin))
basins_clean_3395 <- st_transform(basins_clean, crs=3395) |> 
  select(!area)
basin_bbox <- st_bbox(basins_clean)

# Free-flowing rivers
ffr_in <- "C:\\Users\\user\\Documents\\Articles\\gis_layers\\hydro_data\\sa_FFR_river_v1\\sa_FFR_river_v1.shp"
ffr <- read_sf(ffr_in) |> 
  filter(RIV_ORD <= 5) |> filter(DIS_AV_CMS >= 15)
ffr_1a5_poun_3395 <- st_crop(ffr, basin_bbox) |> st_transform(3395)
# with duplicate REACH_ID.... ????
length(unique(ffr_1a5_poun_3395$REACH_ID))
# Add basins to rivers.
ffr_1a5_poun_3395_basins <- sf::st_intersection(ffr_1a5_poun_3395, 
                                                basins_clean_3395)

plot(basins_clean_3395["BASIN_NAME"])

outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_rivers_v2.gpkg"
st_write(ffr_1a5_poun_3395_basins, dsn = outfile, 
         layer = "ffr_1a5_poun_3395", layer_options = "SPATIAL_INDEX=NO", 
         delete_layer = TRUE, append = TRUE)
st_write(basins_clean_3395, dsn = outfile, 
         layer = "basins_poun_3395", layer_options = "SPATIAL_INDEX=NO", 
         delete_layer = TRUE, append = TRUE)

```


Now convert river lines to points regularly spaced by approximately 1 kilometer.
```{r make-points, eval=FALSE}
# Test and see if can run Orinoco, South and North per basin. 
# Then Amazon by subbasin. Maybe setup with parallel to go quicker.
# 1) Load Free-flowing rivers made previously.
infile <- system.file("vector\\poun_rivers_v2.gpkg", package = "TACAR")
ffr_1a5_poun_3395 <- sf::st_read(infile, layer = "ffr_1a5_poun_3395")

# With 1 to 4 overall length 160,208. Norris et al. 2019 was 215,975. 165 Access, 50 not.
# 76% rivers accessible. 57.2% Accessible and unprotected.
# 354070 km with 1 to 5.
lenth_tot <- as.numeric(sum(st_length(ffr_1a5_poun_3395))) / 1000
#Number of reaches by basin
# table(ffr_1a4_poun_3395$BASIN_NAME)
# Amazon    Coastal North  Coastal South   Orinoco 
# 51871          3728           945          8476 
table(ffr_1a5_poun_3395$BASIN_NAME)
#      Amazon Coastal North Coastal South       Orinoco 
#     111942          8833          2321         19440 
 basin_river_lengths <- ffr_1a5_poun_3395 |> 
   mutate(reach_length_km = as.numeric(units::set_units(sf::st_length(ffr_1a5_poun_3395), km))) |> 
   data.frame() |>
   group_by(BASIN_NAME) |> 
   summarise(river_km = sum(reach_length_km)) |> 
   ungroup()
basin_river_lengths # with 1 to 4
#  BASIN_NAME    river_km
#1 Amazon         128771.
#2 Coastal North    9198.
#3 Coastal South    2515.
#4 Orinoco         19724.
basin_river_lengths
#  BASIN_NAME    river_km
# Amazon         280360.
# Coastal North   21194.
# Coastal South    6507.
# Orinoco         46009.
 # 2) Seperate basins. Makes testing and processing easier.
# If you used parallel this would be much more efficient.
# But the time it would take me to code parallel is actually longer than seperate process.
df_ffr_1a5_poun_3395_south <- ffr_1a5_poun_3395 |> 
  dplyr::filter(BASIN_NAME == "Coastal South") |> data.frame()
df_ffr_1a5_poun_3395_north <- ffr_1a5_poun_3395 |> 
  dplyr::filter(BASIN_NAME == "Coastal North") |> data.frame()
df_ffr_1a5_poun_3395_orinoco <- ffr_1a5_poun_3395 |> 
  dplyr::filter(BASIN_NAME == "Orinoco") |> data.frame()
df_ffr_1a5_poun_3395_amazon <- ffr_1a5_poun_3395 |> 
  dplyr::filter(BASIN_NAME == "Amazon") |> data.frame()
# 3) River lines to points.
# South. n = 2568 on 15/7/2024. 2515 km of RIV_ORD 1 to 4 rivers.
# South. n = 6627 on 18/7/2024. 6507 km of RIV_ORD 1 to 5 rivers.
dfpoints_south <- ddply(df_ffr_1a5_poun_3395_south, .(REACH_ID), 
                        .fun = river_points)
sf_points_south <- st_as_sf(dfpoints_south)
mapview::mapview(sf_points_south)
# North - warning number of items to replace is not a multiple of replacement length
# reach_warn <- df_ffr_1a4_poun_3395_north[3632, "REACH_ID"]
# reach_warn_sf <- df_ffr_1a4_poun_3395_north[3632, ] |> 
#  st_as_sf()
# mapview::mapview(reach_warn_sf)
# Solved - caused by empty point geometry from sample of reach with broken lines.
# North - warning row names were found from a short variable and have been discarded
# cant find where this warning comes from........
# North. n = 9442 on 15 July 2024. 9198 km of RIV_ORD 1 to 4 rivers.
# North. n = 21746 on 18 July 2024. 21194 km of RIV_ORD 1 to 5 rivers.
dfpoints_north <- ddply(df_ffr_1a5_poun_3395_north, .(REACH_ID), 
                        .fun = river_points)
# Orinoco. n = 20515 on 15 July 2024 19724 km of RIV_ORD 1 to 4 rivers.
# Orinoco. n =  47730 on 18 July 2024 46009 km of RIV_ORD 1 to 5 rivers.
dfpoints_orinoco <- ddply(df_ffr_1a5_poun_3395_orinoco, .(REACH_ID), 
                        .fun = river_points)

# Now Amazon. By subbasin to reduce memory demands....
# 17:24 - 18:28ish. 132027 points. 128771 km on 15 July 2024.
# 17:23 - 19:30ish. 287103 points. 280360 km on 18 July, FFR 1 to 5.
dfpoints_amazon <- ddply(df_ffr_1a5_poun_3395_amazon, 
                         .(SUBBASIN_FLAG, REACH_ID), 
                        .fun = river_points)
```

Tidy for export.
```{r export-points, eval=FALSE}
sf_res <- bind_rows(dfpoints_amazon, dfpoints_north, 
                    dfpoints_orinoco, dfpoints_south) |> 
  distinct() |> st_as_sf() 

sf_res$x <- st_coordinates(sf_res)[, 1]
sf_res$y <- st_coordinates(sf_res)[, 2]
# Drop any within 10 meter radius?
sf_res <- sf_res |> 
  mutate(coord_text = paste(round(x, -2), round(y, -2), sep="_"))
sf_res$dupe <- duplicated(sf_res$coord_text)
# This remains close to original length. 
# 160189 points, total river km = 160,208
# 353637 points, total river km = 354,070.
sf_res_out <- sf_res |> 
  filter(!dupe, LENGTH_KM >= .5) |> st_as_sf()

# Export for future use. Avoid GDAL spatial index warning.
#outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points.gpkg"
#st_write(sf_res_out, dsn = outfile, 
#         layer = "ffr_1a4_poun_points_3395", 
#         layer_options = "SPATIAL_INDEX=NO", delete_layer = TRUE, append = TRUE)

outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points_v2.gpkg"
st_write(sf_res_out, dsn = outfile, 
         layer = "ffr_1a5_poun_points_3395", 
         layer_options = "SPATIAL_INDEX=NO", delete_layer = TRUE, append = TRUE)
```

## Accessibility and PAs

Extract accessibility and PAs.
```{r extract-raster, eval=FALSE}
# From Norris et al. 2019.
uain <- "inst/raster/uas.grd"
uas <- rast(uain)
newr <- c(rast("inst/raster/access_new.tif"), rast("inst/raster/cost_distance.tif"))
uas_3395 <- project(uas, "epsg:3395")
pe <- st_as_sf(extract(uas_3395, vect(ffr_1a5_poun_points_3395), bind = TRUE))
pe |> data.frame() |> select(Indigenous, Dist..km.) |> summary()
# check NAs. All  where rivers join oceans north east.
rna <- pe |> filter(is.na(Indigenous)) |> select(REACH_ID)
mapview::mapview(rna)
pe <- pe |> filter(!is.na(Indigenous))
pe <- st_as_sf(extract(newr, vect(pe), bind = TRUE))
pe <- pe |> rename("myuse" = "Use")
pe |> data.frame() |> select(access_new, cost_distance) |> summary() 
# Keep size to less tham 100 MB
pe <- pe |> filter(!is.na(access_new)) |> 
  select(!c(OBJECTID, GOID, NOID, NUOID, NDOID, CON_ID, 
            CONTINENT, x, y, coord_text, dupe, mask, aid))
# Export for future use.
# outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points.gpkg"
#outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\poun_river_points.gpkg"
outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points_v2.gpkg"
st_write(pe, dsn = outfile, 
         layer = "ffr_1a5_poun_points_3395", 
         layer_options = "SPATIAL_INDEX=NO", delete_layer = TRUE, append = TRUE)
```


```{r check-correlations, eval=FALSE}
# Check correlations.
# load points made previously.
# infile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points.gpkg"
# ffr_1a4_poun_points_3395 <- st_read(dsn = infile, 
#                                    layer = "ffr_1a4_poun_points_3395")
infile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points_v2.gpkg"
ffr_1a5_poun_points_3395 <- st_read(dsn = infile, 
                                    layer = "ffr_1a5_poun_points_3395")
ffr_1a5_poun_points_3395 |> 
  data.frame() |> 
  ggplot(aes(x = cost_distance, y = Dist..km.)) + 
  geom_point() + stat_smooth(method = "lm") + 
  coord_equal()
# cor = 0.57, t = 280.11, df = 160167, p-value < 2.2e-16, with ffr_1a4
# cor = 0.56, t = 406.59, df = 353435, p-value < 2.2e-16
cor.test(ffr_1a5_poun_points_3395$cost_distance, ffr_1a5_poun_points_3395$Dist..km.)
  
```

## HydroBasin codes

Add HydroBasin levels 3, 4 and 5. These will also be used to generate 
a flag to exclude any basins where the species does not occur.

HYBAS_ID
Unique basin identifier. The code consists of 10 digits:
• First 1 digit represents the region: 1 = Africa; 2 = Europe; 3 = Siberia; 4 = Asia;
5 = Australia; 6 = South America; 7 = North America; 8 = Arctic (North
America); 9 = Greenland
• Next 2 digits define the Pfafstetter level (01-12). The value ‘00’ is used for the
‘Level 0’ layer that contains all original sub-basins and all Pfafstetter codes (at
all levels); ‘Level 0’ only exists in the standard format of HydroBASINS
(without lakes).
• Next 6 digits represent a unique identifier within the HydroSHEDS network;
values larger than 900,000 represent lakes and only occur in the customized
format (with lakes)
• Last 1 digit indicates the side of a sub-basin in relation to the river network (0 =
noSide; 1 = Left; 2 = Right). Sides are only defined for the customized format
(with lakes).

Pfaf_id
The Pfafstetter code. For general description see literature (e.g., Verdin and Verdin
1999). The Pfafstetter code uses as many digits as the level it represents. This field
can be used to cluster or subdivide sub-basins into nested regions. This field is only
available for levels 1-12 (i.e. not for the ‘Level 0’ layer of the standard format).
```{r join-hydrobasins, eval=FALSE}
# bbox to reduce processing time
poun_bbox_3395 <- sf::st_bbox(ffr_1a5_poun_3395)
# HydroBasin files
hydrobasin_l3_in <- "C:\\Users\\user\\Documents\\Articles\\gis_layers\\hydro_data\\hybas_sa_lev01-12_v1c\\hybas_sa_lev03_v1c.shp"
hydrobasin_l4_in <- "C:\\Users\\user\\Documents\\Articles\\gis_layers\\hydro_data\\hybas_sa_lev01-12_v1c\\hybas_sa_lev04_v1c.shp"
hydrobasin_l5_in <- "C:\\Users\\user\\Documents\\Articles\\gis_layers\\hydro_data\\hybas_sa_lev01-12_v1c\\hybas_sa_lev05_v1c.shp"
# load files
hydrobasins_l3 <- sf::read_sf(hydrobasin_l3_in) |> sf::st_transform(crs=3395) |> 
  sf::st_make_valid() |> sf::st_cast("MULTIPOLYGON") |> 
  sf::st_crop(poun_bbox_3395)
hydrobasins_l4 <- sf::read_sf(hydrobasin_l4_in) |> sf::st_transform(crs=3395) |> 
  sf::st_make_valid() |> sf::st_cast("MULTIPOLYGON") |> 
  sf::st_crop(poun_bbox_3395)
hydrobasins_l5 <- sf::read_sf(hydrobasin_l5_in) |> sf::st_transform(crs=3395) |> 
  sf::st_make_valid() |> sf::st_cast("MULTIPOLYGON") |> 
  sf::st_crop(poun_bbox_3395)

# Points
infile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points_v2.gpkg"
ffr_1a5_poun_points_3395 <- sf::st_read(dsn = infile, 
                                    layer = "ffr_1a5_poun_points_3395")
# appears to be 114 duplicates
ffr_1a5_poun_points_3395 |> data.frame() |>
  dplyr::group_by(reach_id_point) |> 
  dplyr::summarise(acount = dplyr::n()) |> 
  dplyr::filter(acount > 1) |> 
  dplyr::arrange(desc(acount))
# check duplicates. Here from French Guiana, same reach id in Coastal North and South
ffr_1a5_poun_points_3395 |> data.frame() |>
  dplyr::filter(reach_id_point == "60158665_1")
# Here duplicate REACH_ID, both from Amazon, different subbasins - 
# Amazon floodplain and Minor Amazon tribs Pacaj\xe1
ffr_1a5_poun_points_3395 |> data.frame() |>
  dplyr::filter(reach_id_point == "60500077_1")
# Add HydroBasins to rivers. 19:26 - 19:46
#geom column name
geocol <- attr(ffr_1a5_poun_points_3395,"sf_column")

ffr_1a5_poun_3395_hb_l3 <- sf::st_intersection(ffr_1a5_poun_points_3395 |> 
                                                 dplyr::select(REACH_ID, reach_id_point, 
                                                        BASIN_FLAG, SUBBASIN_FLAG), 
                                                hydrobasins_l3 |> 
                                                 dplyr::select(HYBAS_ID, PFAF_ID, SUB_AREA) |> 
                                                 dplyr::rename(c("HYBAS_ID_l3" = "HYBAS_ID", 
                                                               "PFAF_ID_l3" = "PFAF_ID", 
                                                               "SUB_AREA_l3" = "SUB_AREA"))) |>
  data.frame() |> dplyr::select(!tidyselect::all_of(geocol))
ffr_1a5_poun_3395_hb_l4 <- sf::st_intersection(ffr_1a5_poun_points_3395 |> 
                                                 dplyr::select(REACH_ID, reach_id_point, 
                                                        BASIN_FLAG, SUBBASIN_FLAG), 
                                                hydrobasins_l4 |> 
                                                 dplyr::select(HYBAS_ID, PFAF_ID, SUB_AREA)|> 
                                                 dplyr::rename(c("HYBAS_ID_l4" = "HYBAS_ID", 
                                                               "PFAF_ID_l4" = "PFAF_ID", 
                                                               "SUB_AREA_l4" = "SUB_AREA"))) |>
  data.frame() |> dplyr::select(!tidyselect::all_of(geocol))
ffr_1a5_poun_3395_hb_l5 <- sf::st_intersection(ffr_1a5_poun_points_3395 |> 
                                                 dplyr::select(REACH_ID, reach_id_point, 
                                                        BASIN_FLAG, SUBBASIN_FLAG), 
                                                hydrobasins_l5 |> 
                                                 dplyr::select(HYBAS_ID, PFAF_ID, SUB_AREA)) |> 
                                                 dplyr::rename(c("HYBAS_ID_l5" = "HYBAS_ID", 
                                                               "PFAF_ID_l5" = "PFAF_ID", 
                                                               "SUB_AREA_l5" = "SUB_AREA")) |>
  data.frame() |> dplyr::select(!tidyselect::all_of(geocol))

points_hydrobasin <- ffr_1a5_poun_3395_hb_l3 |> 
  dplyr::left_join(ffr_1a5_poun_3395_hb_l4) |> 
  dplyr::left_join(ffr_1a5_poun_3395_hb_l5)

# Now join back with points
ffr_1a5_poun_points_hb_3395 <- ffr_1a5_poun_points_3395 |> 
  dplyr::left_join(points_hydrobasin)
# Export for future use.
outfile <- "C:\\Users\\user\\Documents\\Articles\\2024_Norris_Greenstatus\\TACAR\\inst\\vector\\poun_river_points_v2.gpkg"
# remove some columns to keep filesize below 100 MB.
ffr_1a5_poun_points_hb_3395 |> 
  dplyr::select(!c(UPLAND_SKM, BB_LEN_KM, BB_OCEAN)) |>
sf::st_write(dsn = outfile, 
         layer = "ffr_1a5_poun_points_3395", 
         layer_options = "SPATIAL_INDEX=NO", delete_layer = TRUE, append = TRUE)
sf::st_layers(outfile)
```

